---
title: "Shiny Server Startup Instructions"
author: "Matthew Ross"
date: "December 4, 2015"
output: html_document
---

This is a set of scripts that should help you get your shiny server started. I wrote it with Duke's shiny server setup in mind, but it should be relatively generalization to any Linux server. Most of it comes from Richard Marinos, and I'm just formalizing his very helpful emails. You will have to use Terminal or Command line emulator (like PuTTY) to get started. 

##Getting into the Server From DUke
1. Request a shiny server instance through https://vm-manage.oit.duke.edu/
2. Copy the server address they give you 
3. Go into Terminal (mac) or a command line emulator like PuTTY (windows) and type
  + For windows, make sure SSH type is what is selected to connect
```{r, eval=FALSE}
#Mac
ssh bitnami@servername.oit.duke.edu
#Windows, with SSH
bitnami@servername.oit.duke.edu
```

4. Hit enter, or open.
5. Enter the password given to you by Duke OIT


##Commands once you are in the Linux Shell
Once you have accessed the server you will be working in what is called a "Linux Shell." All this means is that the server is powered by Linux and all the commands to install packages in R or making new directories, will have to be done in this command line. There are lots of tutorials online for working with Linux if you need to do more advanced stuff. 

1. Change your password by 

```{r, eval=FALSE}
passwd NEWPASSWORD
```

2. Install your favorite packages in R as "Root" user, so that they are permanently stored on the server. This is just a wrapper around the classic "install.packages" that you would use inside R. I have a list of key packages here. But you can add more by just adding a comma and then the package name. This will work for most pacages, but all the geospatial ones require installing GDAL on the server which is more complicated, instructions for doing that are in another file in this folder "GDAL Server Install"

```{r, eval=FALSE}
sudo su - root -c "R -e \"install.packages(c('lubridate','ggplot2','shiny','rgl','shinyRGL','sp','raster','rasterVis','reshape2','shape','maptools','fields'),repos='http://cran.rstudio.com/')\""
```

3. Change directory to the shiny server directory
```{r, eval=FALSE}
cd /srv/shiny-server
```

4. Make a new directory for your app
```{r, eval=FALSE}
sudo mkdir APPNAME
```

5. Change ownership from 'root' to bitnami
```{r, eval=FALSE}
sudo chown bitnami APPNAME
```

##Uploading files to app folder
Now is the easier GUI part of the server setup. You can use any FTP client, I prefer cyberduck, to upload your shiny app and its data. 
1. Download and install cyberduck
2. Click "Open Connection"
3. Change type to "SFTP" to connect via ssh
4. Paste the server address should be something like "colab-sbx-65.oit.duke.edu" but with a different number. 
5. Enter username as "bitnami""
6. Enter your password.
7. You should see a bunch of folders, navigate to your shiny app folder which should be in /srv/shiny-server/APPNAME.
8. Drag and drop your files into this server. 
9. **Key** The app will not work if you have directories pointing outside of the folder. All data should be inside the folder and directories should be setup locally only. 
10. Now your app should be up and running, by navigating to port 3838/appname. Like here: http://colab-sbx-65.oit.duke.edu:3838/MTM_Explore/

Please improve this tutorial by adding your thoughts or problems you run into. 


